Stack Safety as a Security Property

### 2021-06-15 submission2/response.txt

> 
> Strengths and Weaknesses
> ------------------------
> I really enjoyed this paper and would
> highly recommend it to anyone working on low-level security
> mechanisms.  The exposition is clear, precise, well-motivated, and
> in many ways synthesizing many years of insights into what it really
> means to protect the call stack.  The examples are well-thought out,
> the definitions unimpeachable, and the clarity of thought
> impressive.  Throughout reading the paper, I anticipated giving a
> 'strong accept' rating.
>
>  What gives me pause is the relative lack of validation or even
> discussion of why there isn't more validation.  I don't know if the
> authors simply haven't had time yet, think it is unnecessary, or ran
> into challenges or complexities that remain unexplained.  I do
> consider significant the property-based testing that was done both
> to validate the simpler tagging approach and identify (obvious in
> hindsight, but that's not a criticism) problems with the lazier
> tagging.  But that still leaves:
>
> * Any discussion on whether it is
> feasible/desirable/in-progress/etc. to prove enforcement mechanisms
> correct.  After all, the authors already have a fully formal
> definition and an implementation in Coq.  While this sort of work
> does not necessarily need a correctness proof in Coq -- I still lean
> slightly toward accept after all -- it seems quite strange not to
> discuss even the possibility of validation including the ability to
> pursue a correctness argument for an enforcement mechanism beyond
> property testing.

We will extend the paper with a more detailed discussion of our roadmap for verification and 
testing, and the issues that we expect to arise.

These properties would absolutely be amenable to a correctness proof. Such a proof would be a
lot of work in a setting with a full ISA, but there is no reason it should be impossible. The
primary challenge we would predict in such an effort is in determining for a given
enforcement mechanism what invariants it needs to maintain in order to work correctly, and proving that
it maintains them. [APT: That last sentence is not a very surprising observation; suggest omitting it.]
This is very feasible for a proof of an individual, well-understood enforcement
mechanism, but not for our goal of evaluating several such mechanisms.  [APT: surely
that depends on our resources.]
[RB: Agreed, it could also be framed in terms of testing as a potential
low effort, high reward strategy.]

@@

We believe that callee-save registers can be described in our model, although that would require tweaks to the domain system.
[APT: for my info, what tweaks?]
They actually provide another use case for observational properties! If we focus on integrity
the reason becomes clear. Non-argument [APT: "callee-save" ??] registers will naturally be overwritten (violating
stepwise integrity) but should always be restored before a return. This behavior corresponds
to marking such registers as permanently sealed (never unsealing them) [APT: ?? don't they get
unsealed after return?] in the context of [APT:??]
observational integrity. (Note that while our specific properties leave registers unsealed,
the formalization permits the same set of domains to apply to them [APT: ??] as to addresses.)

Confidentiality can be applied similarly. We would permanently treat the stack pointer and
argument registers as passed, and so exempt from confidentiality. Then, when a non-argument [APT: "callee-save"??]
register is copied to the stack and later restored before return, the register value does
not represent a visible difference in state between the entry and the return. [APT: I didn't
understand this para]

We had briefly considered this complexity and chosen to avoid it, because it is an extra layer
of detail to apply domains to registers.

[RB: It is worth pointing out that the prior work of Skorstengaard et al. also
relies on a calling convention which clears all non-argument registers before
calls and all non-return registers before returns.]

### 2021-06-15 submission2/response_formal.txt

4. Draw a clearer distinction between language features that are compatible with our
   model but not assumed, those that are required for the model to function,
   and those that are incompatible.

** Detailed Comments **

* Reviewer A *

We will extend the paper with a more detailed discussion of our roadmap for verification and
testing, and the issues that we expect to arise. Here is a rough
outline of how it would cover your concerns.

@@

> ...discussion on whether it is feasible/desirable/in-progress/etc. to prove enforcement mechanisms correct.

These properties would absolutely be amenable to a correctness proof. Such a proof would be a
lot of work in a setting with a full ISA, but there is not reason it should be impossible.
Testing simply makes more sense for our goal of evaluating multiple enforcement mechanisms
to uncover possible bugs -- it is lower effort, but if bugs are found, high reward.

@@

> * The testing for the lazy tagging/clearing found mistakes but the "potential fix [is] itself a prime candidate for future testing" [line 868]
> 
> * Any validation for the lazier observability definition (this is listed as future work.)^M
>
> * Any validation for the coroutine work.

At present we are working to implement
testing of observable properties, and to apply this to the putatively fixed version of the lazy tagging/clearing micro-policy.
Observational integrity testing is implemented, but observational confidentiality testing is not -- since the lazy policy enforces
normal, stepwise confidentiality this is sufficient to test it. [APT: Do we have the results of these tests?] But we did not have this ready before
the deadline. We will therefore take care not
to describe it as a contribution.

Our other immediate goal is to implement testing of a secondary architecture. We will evaluate a Cheri
implementation of stack safety to demonstrate that our properties are not strictly tied to tag-based
enforcement. [APT: Current status of this?]

@@

We consider validation of the passing system to be a secondary priority and coroutines tertiary.
The passing system has existing enforcement mechanisms and serves an important role in common
stack use-cases. Coroutines will require the development of an enforcement mechanism (likely
tag-based) to validate.

> There is one topic beyond my evaluation above worth considering: callee-save registers.
> I fear these may be fairly tricky to support in your set-up.  Upon function entry, callee-save
> registers hold data private to the caller that may be passed arbitrarily deeper into the call stack,
> then saved into arbitrary stack locations, but then must be restored before return.
> Have you or the closely related prior work on enforcement mechanisms considered this complexity?
> The invariants are clear enough informally, but the issue seems to add substantial difficulty to
> tracking that a function faithfully implements the calling convention.

Callee-save registers are plausible in our setup, but would require tweaks to the domain system.
[APT: for my info, what tweaks?] [SNA: I should be clearer -- it needs its own initCtx and updateCtx functions.]
They actually provide another use-case for observational properties! Focusing on integrity:
callee-save registers may be overwritten (violating stepwise integrity) but should always be
restored before a return. This behavior corresponds to marking such registers as permanently sealed
while enforcing observational integrity -- recall that observational integrity compares sealed
components at the start and end of a callee, so the interim changes are not violations.
(Note as well that the formalization permits domains such as "sealed" to apply to registers as well
as to addresses, we just haven't taken advantage of this ability.)

Confidentiality can be applied similarly. We would permanently treat the stack pointer and
argument registers as "passed", and so exempt from confidentiality. Then, when a callee-save
register is copied to the stack and later restored before return, the register value does
not represent a visible difference in state between the entry and the return. [APT: I didn't
understand this para]

@@

In the longer term we would like to extend our model with a heap, and then the interplay between
heap-derived and stack-derived pointers (with similar behaviors but different lifespans) will
be much more interesting.

* Reviewer B *

> Q1. How does this work improve upon Skorstengaard et al? Neither the introduction nor related work section give a satisfying answer to this.^M

We improve upon Skorstengaard et. al. in the following ways.
1. We capture the dynamic nesting of security concerns implicit in stack-based reasoning [APT: what does this mean?]
2. We therefore do not assume a static division of trusted and untrusted code

@@

3. We clearly state the property in terms of a machine's behavior, without appealing to the reader's
understanding of a specific abstract machine [APT: I don't understand this distinction.]
4. Skorstengaard's approach to proving safety via fully-abstract overlay semantics relies on the abstract
machine being architecturally similar to the concrete one (beyond the specific call and return semantics
that distinguish them). Therefore it is less portable: each architecture it applies to wants its own
abstract machine. Our properties apply directly to a wide range of architectures.

We also differ from Skorstengaard et. al. in our focus. We use the language of formal security properties
to apply the established intuitions and formal techniques of non-interference. This demonstrates that stack
safety is expressible as a formal security property.

> Q2. In L117-121 you state certain assumptions about the language semantics. It is unclear to me if these assumptions restrict the languages to which your stack safety definition apply. For example, does a language with a heap allow you to break stack safety? What about a language with delimited continuations that manipulate the stack?^M

The section should be broken into assumptions that are necessarily for our properties, and assumptions
that we explicitly do not make. [APT: ??] Our stack safety model applies regardless of how a language uses non-stack,
non-code memory, including to languages with a heap. They apply to any layout of memory as long as it is
known statically.

### 2021-06-15 submission2/response_formal.txt

These properties would absolutely be amenable to a correctness proof. Such a proof would be a
lot of work in a setting with a full ISA, but there is not reason it should be impossible.
Testing simply makes more sense for our goal of evaluating multiple enforcement mechanisms
to uncover possible bugs -- it is lower effort, but if bugs are found, high reward.

> * The testing for the lazy tagging/clearing found mistakes but the "potential fix [is] itself a prime candidate for future testing" [line 868]
> 
> * Any validation for the lazier observability definition (this is listed as future work.)^M
>
> * Any validation for the coroutine work.

At present we are working to implement
testing of observable properties, and to apply this to the putatively fixed version of the lazy tagging/clearing micro-policy.
Observational integrity testing is implemented, but observational confidentiality testing is not -- since the lazy policy enforces
normal, stepwise confidentiality this is sufficient to test it. [APT: Do we have the results of these tests?] But we did not have this ready before
the deadline. We will therefore take care not
to describe it as a contribution.

Our other immediate goal is to implement testing of a secondary architecture. We will evaluate a Cheri
implementation of stack safety to demonstrate that our properties are not strictly tied to tag-based
enforcement. [APT: Current status of this?]
[RB: SNA refactored the models and Cerise (the capability machine development)
fits quite well in it, so a lot of the machinery for generation and testing is
easily reusable. The tricky part will probably be the generation of blessed
sequences corresponding to the StkTokens-like calling convention, as this is
quite involved.]

### 2021-06-16 submission2/response_formal.txt

* Reviewer A *

We will extend the paper with a more detailed discussion of our roadmap for verification and
testing, and the issues that we expect to arise. Here is a rough
outline of how it would cover your concerns.

> ...discussion on whether it is feasible/desirable/in-progress/etc. to prove enforcement mechanisms correct.

These properties would absolutely be amenable to a correctness proof. Such a
proof would be a lot of work in a setting with a full ISA, but there is no
reason it should be impossible.  We concentrated here on testing because it
is better aligned with our immediate goal of evaluating multiple enforcement
mechanisms to uncover possible bugs -- it is lower effort, but if bugs are
found, high reward.

[BCP: Might be a little dangerous to say "multiple" here because we didn't
actually do very many yet.]

@@

> * The testing for the lazy tagging/clearing found mistakes but the "potential fix [is] itself a prime candidate for future testing" [line 868]
>
> * Any validation for the lazier observability definition (this is listed as future work.)^M
>
> * Any validation for the coroutine work.

At present we are working to implement
testing of observable properties, and to apply this to the putatively fixed version of the lazy tagging/clearing micro-policy.
Observational integrity testing is implemented, but observational confidentiality testing is not -- since the lazy policy enforces
normal, stepwise confidentiality this is sufficient to test it. [APT: Do we have the results of these tests?] But we did not have this ready before
the deadline. We will therefore take care not
to describe it as a contribution.  [BCP: Last sentence not needed.]

Our other immediate goal is to implement testing of a second, rather
different architecture. We plan to evaluate a Cheri implementation of stack
safety to demonstrate that our properties are not strictly tied to tag-based
enforcement. [APT: Current status of this?]  [RB: SNA refactored the models
and Cerise (the capability machine development) fits quite well in it, so a
lot of the machinery for generation and testing is easily reusable. The
tricky part will probably be the generation of blessed sequences
corresponding to the StkTokens-like calling convention, as this is quite
involved.] [BCP: And thus likely to be wrong. :-) ]

We consider validation of the passing system [BCP: ??] to be a secondary
priority and coroutines tertiary.  The passing system has existing
enforcement mechanisms and serves an important role in common stack
use-cases. Coroutines will require the development of an enforcement
mechanism (e.g., one based on tags) to validate.

@@

> There is one topic beyond my evaluation above worth considering: callee-save registers.
> I fear these may be fairly tricky to support in your set-up.  Upon function entry, callee-save
> registers hold data private to the caller that may be passed arbitrarily deeper into the call stack,
> then saved into arbitrary stack locations, but then must be restored before return.
> Have you or the closely related prior work on enforcement mechanisms considered this complexity?
> The invariants are clear enough informally, but the issue seems to add substantial difficulty to
> tracking that a function faithfully implements the calling convention.

Callee-save registers are plausible in our setup, but would require tweaks to the domain system.
[APT: for my info, what tweaks?] [SNA: I should be clearer -- it needs its own initCtx and updateCtx functions.]
They actually provide another use-case for observational properties! Focusing on integrity:
callee-save registers may be overwritten (violating stepwise integrity) but should always be
restored before a return. This behavior corresponds to marking such registers as permanently sealed
while enforcing observational integrity -- recall that observational integrity compares sealed
components at the start and end of a callee, so the interim changes are not violations.
(Note as well that the formalization permits domains such as "sealed" to apply to registers as well
as to addresses, we just haven't taken advantage of this ability.)

Confidentiality can be applied similarly. We would permanently treat the stack pointer and
argument registers as "passed", and so exempt from confidentiality. Then, when a callee-save
register is copied to the stack and later restored before return, the register value does
not represent a visible difference in state between the entry and the return. [APT: I didn't
understand this para]

We had briefly considered adding this complexity and chosen to avoid it, because it is an extra layer
of detail to apply domains to registers.

[RB: It is worth pointing out that the prior work of Skorstengaard et al. also
relies on a calling convention which clears all non-argument registers before
calls and all non-return registers before returns.] [BCP: Yes!]

@@

> Another minor question: The difference between [nested] call-by-reference and passing
> address-of-local was interesting, but I'm not sure how much it makes sense to argue the
> latter is "beyond the scope of stack safety per se" [line 939] without arguing the same
> of the former.  Perhaps the argument can be made that address-of-local can be put in the
> heap while a by-reference parameter cannot and since you don't guarantee heap safety,
> you shouldn't try to do anything useful with address-of-local?

Yes, this is the sort of issue that makes address-of-local hard to handle without heap safety.
More broadly the issue is one of assuming intent on the part of the code. We tend to
assume that an address-of-local should be treated similarly to the address of a heap object,
but with a different lifetime. For instance, if a heap model treats heap pointers like capabilities,
we would treat stack pointers the same. In heap model that permits pointer forging (restricted to
within compartments, say,) we would expect to be able to forge pointers to address-taken locals
in the same circumstances. By contrast, the expected behavior of stack allocated arguments is
closely tied to that of the stack itself.

In the longer term we would like to extend our model with a heap,[BCP: Cite
Arthur's paper and point out that it is phrased in similar terms and should
thus be compatible!] and then the interplay between heap-derived and
stack-derived pointers (with similar behaviors but different lifespans) will
be much more interesting.

@@

* Reviewer B *

> Q1. How does this work improve upon Skorstengaard et al? Neither the introduction nor related work section give a satisfying answer to this.^M

We improve upon Skorstengaard et. al. in the following ways.
1. We capture the dynamic nesting of security concerns implicit in stack-based reasoning [APT: what does this mean?]
2. We therefore do not assume a static division of trusted and untrusted code

Taken together, these distinctions give the practical advantage that we can reason about stack safety
in contexts where a buggy function is compromised and used to attack other
instances of the same function.  [BCP: Example? Plus a clear explanation of
why S. et al cannot handle it while we can.]

3. We clearly state the property in terms of a machine's behavior, without appealing to the reader's
understanding of a specific abstract machine [APT: I don't understand this distinction.]
4. Skorstengaard's approach to proving safety via fully-abstract overlay semantics relies on the abstract
machine being architecturally similar to the concrete one (beyond the specific call and return semantics
that distinguish them). Therefore it is less portable: each architecture it applies to wants its own
abstract machine. Our properties should apply directly to a wide range of
architectures.

@@

We want our properties to be applicable even to low-level code that may not keep function
footprints isolated from one another. Consider a program that mixes C code with hand-written
assembly -- the assembly may make calls and returns, which should be protected by stack safety,
as well as arbitrary jumps, which should not.

In Skorstengaard, an untrusted adversary can exhibit such behavior, but only trusted code can
expect its data to be protected. We extend this protection to all callers, even those that
exhibit unusual control flow. [APT: Don't think I get this.]

* Reviewer C *

> - It is a bit difficult to assess the merits of this work. Yes, the
>   stack safety property is important for security, but what new results
>   or new capabilities does the proposed definition bring in? If it is
>   positioned as a security solution, does it lead to more efficient
>   defenses of control flow attacks? Does it prevent any new attacks
>   that existing defenses fail? The paper does not talk about data flow
>   across heap and stack, which is an important element for many
>   real-world memory corruption attacks.

[BCP: We definitely need to respond directly to this.  First, we should
agree that the title is confusing (of course it is a security property) and
needs changing, and then we should say clearly what we meant by it.]

@@

> I might have missed something important. Are there any applications of the proposed formalization?

A formal definition of a type of security is valuable in its own right. The main application
is identifying bugs in real world enforcement mechanisms -- without such a definition, we
can't necessarily agree that a given behavior is a bug. It gives us the ability to reason
clearly about what is gained from applying a given enforcement mechanism to a system. Since
many enforcement mechanisms are quite costly in speed, dollars, and silicon, it is important
to understand precisely what they buy us.

[BCP: Also, we intended to use it for bug finding, and the case study
confirms that it is suitable for this purpose.]

### 2021-06-16 submission2/response_formal.txt

These properties would absolutely be amenable to a correctness proof. Such a proof would be a
lot of work in a setting with a full ISA, but there is not reason it should be impossible.
Testing simply makes more sense for our goal of evaluating multiple enforcement mechanisms
to uncover possible bugs -- it is lower effort, but if bugs are found, high reward.

> * The testing for the lazy tagging/clearing found mistakes but the "potential fix [is] itself a prime candidate for future testing" [line 868]
> 
> * Any validation for the lazier observability definition (this is listed as future work.)^M
>
> * Any validation for the coroutine work.

At present we are working to implement
testing of observable properties, and to apply this to the putatively fixed version of the lazy tagging/clearing micro-policy.
Observational integrity testing is implemented, but observational confidentiality testing is not -- since the lazy policy enforces
normal, stepwise confidentiality this is sufficient to test it. [APT: Do we have the results of these tests?] But we did not have this ready before
the deadline. We will therefore take care not
to describe it as a contribution.

Our other immediate goal is to implement testing of a secondary architecture. We will evaluate a Cheri
implementation of stack safety to demonstrate that our properties are not strictly tied to tag-based
enforcement. We are currently working on the QuickChick generators for Cheri programs, which have a
different (more complicated) calling convention.

@@

We consider validation of the passing system to be a secondary priority and coroutines tertiary.
The passing system has existing enforcement mechanisms and serves an important role in common
stack use-cases. Coroutines will require the development of an enforcement mechanism (likely
tag-based) to validate.

> There is one topic beyond my evaluation above worth considering: callee-save registers.
> I fear these may be fairly tricky to support in your set-up.  Upon function entry, callee-save
> registers hold data private to the caller that may be passed arbitrarily deeper into the call stack,
> then saved into arbitrary stack locations, but then must be restored before return.
> Have you or the closely related prior work on enforcement mechanisms considered this complexity?
> The invariants are clear enough informally, but the issue seems to add substantial difficulty to
> tracking that a function faithfully implements the calling convention.

Callee-save registers are plausible in our setup, using a version of the sharing model
with its context initialization and updates modified to apply appropriate domains to registers.
They actually provide another use-case for observational properties! Focusing on integrity:
callee-save registers may be overwritten (violating stepwise integrity) but should always be
restored before a return. This requirement corresponds to observational integrity with the
callee-save registers sealed -- recall that observational integrity compares sealed
components at the start and end of a callee, so the interim changes are not violations.

Confidentiality can be applied similarly. A callee-save register should be secret, but
in observational confidentiality, if it is copied to the stack and then restored before
return, the property will allow the original register to be read normally after return.
But the property will not allow it to be leaked in any other way.

@@

In the longer term we would like to extend our model with a heap, and then the interplay between
heap-derived and stack-derived pointers (with similar behaviors but different lifespans) will
be much more interesting.

* Reviewer B *

> Q1. How does this work improve upon Skorstengaard et al? Neither the introduction nor related work section give a satisfying answer to this.^M

We improve upon Skorstengaard et. al. in the following ways.
1. We capture a desireable security characteristic of dynamically nested functions, that every caller is safe from its callee
2. We therefore do not assume a static division of trusted and untrusted code

@@

3. Skorstengaard's approach to proving safety via fully-abstract overlay semantics relies on the abstract
machine being architecturally similar to the concrete one (beyond the specific call and return semantics
that distinguish them). Therefore it is less portable: each architecture it applies to wants its own
abstract machine. Our properties apply directly to a wide range of architectures.

We also differ from Skorstengaard et. al. in our focus. We use the language of formal security properties
to apply the established intuitions and formal techniques of non-interference. This demonstrates that stack
safety is expressible as a formal security property.

> Q2. In L117-121 you state certain assumptions about the language semantics. It is unclear to me if these assumptions restrict the languages to which your stack safety definition apply. For example, does a language with a heap allow you to break stack safety? What about a language with delimited continuations that manipulate the stack?^M

We will edit section 2 to draw a clearer distinction between language features
that are compatible with our model but not assumed, those that are required for
the model to function, and those that are incompatible.

Our stack safety model applies regardless of how a language uses non-stack,
non-code memory, including to languages with a heap. They apply to any layout of memory as long as it is
known statically.

### 2021-06-16 submission2/response_formal.txt

These properties would absolutely be amenable to a correctness proof. Such a
proof would be a lot of work in a setting with a full ISA, but there is no
reason it should be impossible.  We concentrated here on testing because it
is better aligned with our immediate goal of evaluating multiple enforcement
mechanisms to uncover possible bugs -- it is lower effort, but if bugs are
found, high reward.

[BCP: Might be a little dangerous to say "multiple" here because we didn't
actually do very many yet.]

> * The testing for the lazy tagging/clearing found mistakes but the "potential fix [is] itself a prime candidate for future testing" [line 868]
>
> * Any validation for the lazier observability definition (this is listed as future work.)^M
>
> * Any validation for the coroutine work.

At present we are working to implement
testing of observable properties, and to apply this to the putatively fixed version of the lazy tagging/clearing micro-policy.
Observational integrity testing is implemented, but observational confidentiality testing is not -- since the lazy policy enforces
normal, stepwise confidentiality this is sufficient to test it. [APT: Do we have the results of these tests?] But we did not have this ready before
the deadline.

Our other immediate goal is to implement testing of a secondary architecture. We will evaluate a Cheri
implementation of stack safety to demonstrate that our properties are not strictly tied to tag-based
enforcement. We are currently working on the QuickChick generators for Cheri programs, which have a
different calling convention -- a more complicated one, which is easier to get wrong.

We consider validation of the passing system [BCP: ??] to be a secondary
priority and coroutines tertiary.  The passing system has existing
enforcement mechanisms and serves an important role in common stack
use-cases. Coroutines will require the development of an enforcement
mechanism (e.g., one based on tags) to validate.

@@

* Reviewer B *

> Q1. How does this work improve upon Skorstengaard et al? Neither the introduction nor related work section give a satisfying answer to this.^M

We improve upon Skorstengaard et. al. in the following ways.
1. We capture a desireable security characteristic of dynamically nested functions, that every caller is safe from its callee
2. We therefore do not assume a static division of trusted and untrusted code

Taken together, these distinctions give the practical advantage that we can reason about stack safety
in contexts where a buggy function is compromised and used to attack other
instances of the same function.  [BCP: Example? Plus a clear explanation of
why S. et al cannot handle it while we can.]

3. Skorstengaard's approach to proving safety via fully-abstract overlay semantics relies on the abstract
machine being architecturally similar to the concrete one (beyond the specific call and return semantics
that distinguish them). Therefore it is less portable: each architecture it applies to wants its own
abstract machine. Our properties should apply directly to a wide range of
architectures.

### 2021-06-16 submission2/response_formal.txt

[BCP: Might be a little dangerous to say "multiple" here because we didn't
actually do very many yet.]

> * The testing for the lazy tagging/clearing found mistakes but the "potential fix [is] itself a prime candidate for future testing" [line 868]
>
> * Any validation for the lazier observability definition (this is listed as future work.)^M
>
> * Any validation for the coroutine work.

At present we are working to implement
testing of observable properties, and to apply this to the putatively fixed version of the lazy tagging/clearing micro-policy.
Observational integrity testing is implemented, but observational confidentiality testing is not -- since the lazy policy enforces
normal, stepwise confidentiality this is sufficient to test it. [APT: Do we have the results of these tests?] But we did not have this ready before
the deadline.

Our other immediate priority is to implement testing of a secondary enforcement mechanism. We will evaluate a Cheri
implementation of stack safety to demonstrate that our properties are not strictly tied to tag-based
enforcement. We are currently working on the QuickChick generators for Cheri programs, which have a
different calling convention -- a more complicated one, which is easier to get wrong.

Validation of the argument-passing and coroutining models are longer-term priorities.
The passing system has existing enforcement mechanisms and serves an important role in common stack
use-cases. Coroutines will require the development of an enforcement
mechanism (e.g., one based on tags) to validate.

@@

We had briefly considered adding this complexity and chosen to avoid it, because it is an extra layer
of detail to apply domains to registers.

[RB: It is worth pointing out that the prior work of Skorstengaard et al. also
relies on a calling convention which clears all non-argument registers before
calls and all non-return registers before returns.] [BCP: Yes!]

> Another minor question: The difference between [nested] call-by-reference and passing
> address-of-local was interesting, but I'm not sure how much it makes sense to argue the
> latter is "beyond the scope of stack safety per se" [line 939] without arguing the same
> of the former.  Perhaps the argument can be made that address-of-local can be put in the
> heap while a by-reference parameter cannot and since you don't guarantee heap safety,
> you shouldn't try to do anything useful with address-of-local?

Yes, this is exactly the sort of issue that makes address-of-local hard to handle without heap safety.
More broadly the issue is one of assuming intent on the part of the code. We tend to
assume that an address-of-local should be treated similarly to the address of a heap object,
but with a different lifetime. For instance, if a heap model treats heap pointers like capabilities,
we would treat stack pointers the same. In heap model that permits pointer forging (restricted to
within compartments, say,) we would expect to be able to forge pointers to address-taken locals
in the same circumstances. By contrast, the expected behavior of stack allocated arguments is
closely tied to that of the stack itself.

### 2021-06-16 submission2/response_formal.txt

** Overview **

These reviews have identified some areas that the paper failed to explain adequately.
Fortunately none of these issues are in the formalism itself, but meta questions: does
the work improve on the competition? Do we have a plan to validate the model further?
Can it be extended to other topics? We have answers to these questions, and we can
address them explicitly in the paper.

### 2021-06-16 submission2/response_formal.txt

** Overview **

We thank the reviewers for their comments. The reviews identify 
some areas that the paper fails to explain adequately.
Fortunately [APT: hmm] none of these issues are in the formalism itself, but meta questions: How does
the work improve on existing formalizations of stack safety? Do we have a plan to validate the model further?
Can our work be extended to cover more complex uses of the stack? We have answers to these questions, and we can
address them explicitly in the paper.

@@

** Changes **

1. We will give a more thorough discussion of validation, discuss the feasibility of proofs
   and of randomized testing of more complex properties, and detail our roadmap for
   further validation of different versions of the property.

2. We will give an explicit, direct comparison to the work of Skorstengaard et. al. Currently
   we discuss it briefly in our related work section, but our novelty has not been
   obvious, so we will be more thorough in explaining the differences.

3. We will improve the presentation as suggested. We need to gradually
   introduce the examples and associated intuition (including our
   visual metaphors) as early as possible, starting in the
   introduction.

4. We will more clearly distinguish language features that are compatible with our
   model but not assumed [APT: don't understand this phrasing], those that are required for the model to function,
   and those that are incompatible.

@@

* Reviewer A *

We will extend the paper with a more detailed discussion of our roadmap for verification and
testing, and the issues that we expect to arise. Here is a rough
outline of how it would cover your concerns.

> ...discussion on whether it is feasible/desirable/in-progress/etc. to prove enforcement mechanisms correct.

These properties would absolutely be amenable to a correctness proof. Such a
proof would be a lot of work in a setting with a full ISA, but we see no
fundamental barriers.  We chose to concentrate on testing because it
is better aligned with our immediate goal of evaluating multiple enforcement
mechanisms to uncover possible bugs -- it is lower effort, but if bugs are
found, high reward.

@@

Our other immediate priority is to implement testing of a secondary enforcement mechanism. We will evaluate a Cheri
implementation of stack safety to demonstrate that our properties are not strictly tied to tag-based
enforcement. We are currently working on the QuickChick generators for Cheri programs, which have a
different calling convention -- a more complicated one, which is easier to get wrong.

Validation of the argument-passing and coroutining models are longer-term priorities.
The passing system has existing enforcement mechanisms and serves an important role in common stack
use-cases. Coroutines will require the development of an enforcement
mechanism (e.g., one based on tags) to validate.

> There is one topic beyond my evaluation above worth considering: callee-save registers.
> I fear these may be fairly tricky to support in your set-up.  Upon function entry, callee-save
> registers hold data private to the caller that may be passed arbitrarily deeper into the call stack,
> then saved into arbitrary stack locations, but then must be restored before return.
> Have you or the closely related prior work on enforcement mechanisms considered this complexity?
> The invariants are clear enough informally, but the issue seems to add substantial difficulty to
> tracking that a function faithfully implements the calling convention.

Callee-save registers are compatible with our approach, using a version of the sharing model
with its context initialization and updates modified to apply appropriate domains to registers.
They actually provide another use-case for observational properties! Focusing on integrity:
callee-save registers may be overwritten (violating stepwise integrity) but should always be
restored before a return. This requirement corresponds to observational integrity with the
callee-save registers sealed -- recall that observational integrity compares sealed
components at the start and end of a callee, so the interim changes are not violations.

@@

We had briefly considered adding this complexity and chosen to avoid it, because it is an extra layer
of detail to apply domains to registers.

[RB: It is worth pointing out that the prior work of Skorstengaard et al. also
relies on a calling convention which clears all non-argument registers before
calls and all non-return registers before returns.] [BCP: Yes!]

> Another minor question: The difference between [nested] call-by-reference and passing
> address-of-local was interesting, but I'm not sure how much it makes sense to argue the
> latter is "beyond the scope of stack safety per se" [line 939] without arguing the same
> of the former.  Perhaps the argument can be made that address-of-local can be put in the
> heap while a by-reference parameter cannot and since you don't guarantee heap safety,
> you shouldn't try to do anything useful with address-of-local?

Yes, this was exactly our thinking.  
We know what constitutes legal access to stack-allocated arguments, but for explicit pointers
this is much less clear, and depends on the intent of the programmer. We believe
an address-of-local should be treated similarly to the address of a heap object,
but with a different lifetime. For instance, if a heap safety model treats heap pointers like capabilities
(possession implies access rights), we would treat stack pointers the same.
Similarly, if the heap model permits pointer forging (restricted to
within compartments, say) we would expect to be able to forge pointers to address-taken locals
in the same circumstances.

@@

3. Skorstengaard's approach to proving safety via fully-abstract overlay semantics relies on the abstract
machine being architecturally similar to the concrete one (beyond the specific call and return semantics
that distinguish them). Therefore it is less portable: each architecture it applies to wants its own
abstract machine. Our properties should apply directly to a wide range of
architectures.

We also differ from Skorstengaard et. al. in our focus. We use the language of formal security properties
to apply the established intuitions and formal techniques of non-interference. This demonstrates that stack
safety is expressible as a formal security property.

> Q2. In L117-121 you state certain assumptions about the language semantics. It is unclear to me if these assumptions restrict the languages to which your stack safety definition apply. For example, does a language with a heap allow you to break stack safety? What about a language with delimited continuations that manipulate the stack?^M

We will edit section 2 to draw a clearer distinction between language features
that are compatible with our model but not assumed [APT: again, don't understand this language], those that are required for
the model to function, and those that are incompatible.

Our stack safety model applies regardless of how a language uses non-stack,
non-code memory, so it works for languages with a heap. It applies to any layout of memory as long as that is
known statically.
